---
title: "Sensitivity analysis"
author: "Sumit Jyoti"
date: "2026-08-01"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

rm(list = ls())

library(ggplot2)

dir.create("results", showWarnings = FALSE)
dir.create("images", showWarnings = FALSE)
```

# Read simulation data

```{r read-data}
main_data <- readRDS(
  "SD-DRM output for sensitivity analysis.rds"
)
```

# Select the overall data

The overall analysis includes all contaminated and uncontaminated simulation iterations.

```{r overall-data}
overall_data <- main_data
```

# Select the contaminated-event data

The conditional analysis includes only iterations in which the retail bacterial concentration was greater than zero.

```{r contaminated-data}
contaminated_data <- main_data[
  main_data$C_retail > 0,
]
```

# Analysis sample sizes

```{r sample-sizes}
sample_sizes <- data.frame(
  Analysis = c(
    "Overall",
    "Contaminated exposure events"
  ),
  Number_of_iterations = c(
    nrow(overall_data),
    nrow(contaminated_data)
  )
)

print(sample_sizes)
```

# Spearman sensitivity analysis

```{r spearman-function}
calculate_spearman <- function(
  data,
  input_variables,
  outcome_variable,
  scenario_name,
  analysis_name
) {
  
  data.frame(
    Analysis = analysis_name,
    Scenario = scenario_name,
    Parameter = input_variables,
    
    Spearman_coefficient = sapply(
      input_variables,
      function(input_name) {
        cor(
          data[[input_name]],
          data[[outcome_variable]],
          method = "spearman"
        )
      }
    ),
    
    N = nrow(data),
    row.names = NULL
  )
}
```

# Run the sensitivity analysis

```{r run-analysis}
run_sensitivity_analysis <- function(
  data,
  analysis_name
) {
  
  rbind(
    
    calculate_spearman(
      data = data,
      
      input_variables = c(
        "C_retail",
        "M",
        "F_pathogenic",
        "fr",
        "pMIC",
        "Emax",
        "EC50"
      ),
      
      outcome_variable = "risk_raw",
      scenario_name = "Raw shrimp consumption",
      analysis_name = analysis_name
    ),
    
    calculate_spearman(
      data = data,
      
      input_variables = c(
        "C_retail",
        "M",
        "F_pathogenic",
        "R_under",
        "fr",
        "pMIC",
        "Emax",
        "EC50"
      ),
      
      outcome_variable = "risk_undercooked",
      scenario_name = "Undercooked shrimp consumption",
      analysis_name = analysis_name
    ),
    
    calculate_spearman(
      data = data,
      
      input_variables = c(
        "C_retail",
        "M_h",
        "F_pathogenic",
        "T_shrimp_hand",
        "T_hand_food",
        "fr",
        "pMIC",
        "Emax",
        "EC50"
      ),
      
      outcome_variable = "risk_hand",
      scenario_name = "Hand-mediated cross-contamination",
      analysis_name = analysis_name
    )
  )
}

overall_results <- run_sensitivity_analysis(
  data = overall_data,
  analysis_name = "Overall"
)

contaminated_results <- run_sensitivity_analysis(
  data = contaminated_data,
  analysis_name = "Contaminated exposure events"
)

spearman_results <- rbind(
  overall_results,
  contaminated_results
)
```

# Add parameter labels

```{r parameter-labels}
parameter_labels <- c(
  C_retail = "Retail concentration",
  M = "Shrimp consumption amount",
  M_h = "Raw shrimp mass handled",
  F_pathogenic = "Pathogenic fraction",
  R_under = "Cooking log reduction",
  T_shrimp_hand = "Shrimp-to-hand transfer",
  T_hand_food = "Hand-to-food transfer",
  fr = "Resistant fraction",
  pMIC = "Doxycycline concentration (%MIC)",
  Emax = "Maximum killing effect (Emax)",
  EC50 = "Half-maximal concentration (EC50)"
)

spearman_results$Parameter_label <- unname(
  parameter_labels[
    spearman_results$Parameter
  ]
)

spearman_results$Absolute_coefficient <- abs(
  spearman_results$Spearman_coefficient
)

scenario_order <- c(
  "Raw shrimp consumption",
  "Undercooked shrimp consumption",
  "Hand-mediated cross-contamination"
)

analysis_order <- c(
  "Overall",
  "Contaminated exposure events"
)

spearman_results$Scenario <- factor(
  spearman_results$Scenario,
  levels = scenario_order
)

spearman_results$Analysis <- factor(
  spearman_results$Analysis,
  levels = analysis_order
)
```

# Create a common parameter order

The same parameter order is used in the overall and contaminated-event columns.

```{r parameter-order}
spearman_results$Parameter_plot <- paste(
  spearman_results$Scenario,
  spearman_results$Parameter_label,
  sep = "___"
)

ordering_data <- aggregate(
  Absolute_coefficient ~ Scenario + Parameter_plot,
  data = spearman_results,
  FUN = mean
)

parameter_levels <- unlist(
  lapply(
    scenario_order,
    function(current_scenario) {
      
      scenario_data <- ordering_data[
        as.character(ordering_data$Scenario) ==
          current_scenario,
      ]
      
      scenario_data$Parameter_plot[
        order(
          scenario_data$Absolute_coefficient
        )
      ]
    }
  ),
  use.names = FALSE
)

spearman_results$Parameter_plot <- factor(
  spearman_results$Parameter_plot,
  levels = parameter_levels
)

print(spearman_results)
```

# Save results

```{r save-results}
write.csv(
  spearman_results,
  file.path(
    "results",
    "spearman_sensitivity_results.csv"
  ),
  row.names = FALSE
)
```

# Two-panel sensitivity figure

The parameter labels are shown once on the left and are shared by both analysis columns.

```{r sensitivity-figure}
bar_colour <- "#7895B2"

analysis_labels <- c(
  "Overall" = paste0(
    "A. Overall analysis\n(n = ",
    format(
      nrow(overall_data),
      big.mark = ","
    ),
    ")"
  ),
  
  "Contaminated exposure events" = paste0(
    "B. Contaminated exposure events\n(n = ",
    format(
      nrow(contaminated_data),
      big.mark = ","
    ),
    ")"
  )
)

combined_sensitivity_plot <- ggplot(
  spearman_results,
  aes(
    x = Spearman_coefficient,
    y = Parameter_plot
  )
) +
  geom_col(
    fill = bar_colour,
    colour = "black",
    linewidth = 0.35,
    width = 0.72
  ) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    linewidth = 0.45
  ) +
  geom_text(
    aes(
      label = sprintf(
        "%.2f",
        Spearman_coefficient
      ),
      
      hjust = ifelse(
        Spearman_coefficient >= 0,
        -0.15,
        1.15
      )
    ),
    size = 3
  ) +
  facet_grid(
    rows = vars(Scenario),
    cols = vars(Analysis),
    scales = "free_y",
    space = "free_y",
    
    labeller = labeller(
      Analysis = analysis_labels
    )
  ) +
  scale_y_discrete(
    labels = function(x) {
      sub(
        "^.*___",
        "",
        x
      )
    }
  ) +
  scale_x_continuous(
    breaks = seq(
      -1,
      1,
      by = 0.25
    )
  ) +
  coord_cartesian(
    xlim = c(
      -1.08,
      1.08
    ),
    clip = "off"
  ) +
  labs(
    x = "Spearman rank correlation coefficient",
    y = "Model input"
  ) +
  theme_bw(
    base_size = 11
  ) +
  theme(
    strip.background = element_rect(
      fill = "grey90",
      colour = "grey35"
    ),
    
    strip.text.x = element_text(
      face = "bold",
      size = 11
    ),
    
    strip.text.y = element_text(
      face = "bold",
      size = 10
    ),
    
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    
    axis.title = element_text(
      size = 11
    ),
    
    axis.text = element_text(
      size = 9
    ),
    
    panel.spacing.x = grid::unit(
      0.8,
      "lines"
    ),
    
    panel.spacing.y = grid::unit(
      0.6,
      "lines"
    ),
    
    plot.margin = margin(
      8,
      18,
      8,
      8
    )
  )

print(combined_sensitivity_plot)

ggsave(
  filename = file.path(
    "images",
    "spearman_sensitivity_two_panel.png"
  ),
  plot = combined_sensitivity_plot,
  width = 13,
  height = 9,
  dpi = 300,
  bg = "white"
)
```

# Output files

```text
results/spearman_sensitivity_results.csv

images/spearman_sensitivity_two_panel.png
```