---
title: "Shrimp SD-DRM Using Empirical Emax and EC50 Distributions"
author: "Sumit Jyoti"
date: "2026-07-31"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

rm(list = ls())

library(ggplot2)
library(scales)

source("model_functions.R")
source("Graph_funtion.R")

annual_seed <- 123
set.seed(annual_seed)

dir.create("results", showWarnings = FALSE)
dir.create("images", showWarnings = FALSE)
```

# Read input data

```{r read-input-data}
dose_data <- read.csv(
  "full_simulation_data_three_scenarios.csv"
)

pd_parameters <- read.csv(
  "Emax and EC50.csv",
  check.names = FALSE
)

scenario_list <- list(
  "Raw shrimp consumption" =
    dose_data$D_raw,
  
  "Undercooked shrimp consumption" =
    dose_data$D_under,
  
  "Hand-mediated cross-contamination" =
    dose_data$D_hand
)

scenario_levels <- names(
  scenario_list
)
```

# SD-DRM parameters

```{r sd-drm-parameters}
alpha <- 0.60
beta <- 1.31e6
tfs <- 1
MIC <- 0.13

n_person_years <- 100000

consumption_mean <- 0.0369
consumption_se <- 0.0031
days_per_year <- 365

fr_fixed <- 0.20
pMIC_fixed <- 2

pMIC_seq <- seq(
  0,
  10,
  by = 0.1
)

pMIC_plot <- c(
  0,
  1,
  2,
  5
)

pMIC_breaks <- seq(
  0,
  10,
  by = 1
)

fr_seq <- seq(
  0,
  0.5,
  by = 0.05
)

fr_plot <- c(
  0,
  0.2,
  0.5
)

fr_breaks <- seq(
  0,
  0.5,
  by = 0.05
)
```

# Draw paired Emax and EC50 values

```{r draw-empirical-parameters}
set.seed(annual_seed + 1)

pd_id <- sample(
  seq_len(
    nrow(pd_parameters)
  ),
  size = nrow(dose_data),
  replace = TRUE
)

Emax_mc <- pd_parameters$Emax_per_day[
  pd_id
]

EC50_mc <- pd_parameters$EC50_ug_mL[
  pd_id
]
```

Sampling the row index ensures that each \(E_{\max}\) value remains paired with the \(EC_{50}\) value obtained from the same bootstrap run.

# Save sensitivity-analysis simulation data

The sensitivity dataset contains the original exposure inputs, variable SD-DRM inputs, paired pharmacodynamic parameters, and the corresponding event-level infection risks.

```{r save-sensitivity-data}
set.seed(annual_seed + 2)

n_sensitivity <- nrow(
  dose_data
)

fr_sensitivity <- runif(
  n = n_sensitivity,
  min = min(fr_seq),
  max = max(fr_seq)
)

pMIC_sensitivity <- sample(
  pMIC_seq,
  size = n_sensitivity,
  replace = TRUE
)

C_sensitivity <- (
  pMIC_sensitivity / 100
) * MIC

alpha_s_sensitivity <- numeric(
  n_sensitivity
)

beta_s_sensitivity <- numeric(
  n_sensitivity
)

for (pm in unique(pMIC_sensitivity)) {
  
  index <- which(
    pMIC_sensitivity == pm
  )
  
  susceptible_parameters <- get_ab_beta(
    alpha = alpha,
    beta = beta,
    C = (pm / 100) * MIC,
    Emax = Emax_mc[index],
    EC50 = EC50_mc[index],
    tfs = tfs
  )
  
  alpha_s_sensitivity[index] <- unname(
    susceptible_parameters[["alpha_s"]]
  )
  
  beta_s_sensitivity[index] <- unname(
    susceptible_parameters[["beta_s"]]
  )
}

calculate_sensitivity_risk <- function(dose) {
  
  susceptible_dose <- dose * (
    1 - fr_sensitivity
  )
  
  resistant_dose <- dose *
    fr_sensitivity
  
  p_ext_s <- (
    1 +
      susceptible_dose /
      beta_s_sensitivity
  )^(
    -alpha_s_sensitivity
  )
  
  p_ext_r <- (
    1 +
      resistant_dose /
      beta
  )^(
    -alpha
  )
  
  pmin(
    pmax(
      1 - p_ext_s * p_ext_r,
      0
    ),
    1
  )
}

sensitivity_data <- cbind(
  simulation_id =
    seq_len(n_sensitivity),
  
  dose_data,
  
  pd_id =
    pd_id,
  
  Bootstrap_run =
    pd_parameters$Bootstrap_run[pd_id],
  
  Emax =
    Emax_mc,
  
  EC50 =
    EC50_mc,
  
  fr =
    fr_sensitivity,
  
  pMIC =
    pMIC_sensitivity,
  
  C_doxycycline =
    C_sensitivity,
  
  risk_raw =
    calculate_sensitivity_risk(
      dose_data$D_raw
    ),
  
  risk_undercooked =
    calculate_sensitivity_risk(
      dose_data$D_under
    ),
  
  risk_hand =
    calculate_sensitivity_risk(
      dose_data$D_hand
    )
)

saveRDS(
  sensitivity_data,
  file = "SD-DRM output for sensitivity analysis.rds",
  compress = FALSE
)

rm(
  alpha_s_sensitivity,
  beta_s_sensitivity,
  C_sensitivity,
  fr_sensitivity,
  pMIC_sensitivity,
  sensitivity_data
)
```

# Empirical parameter functions

```{r empirical-functions}
annualise_risk <- function(
  p,
  K_j
) {
  
  K_distribution <- prop.table(
    table(K_j)
  )
  
  K_values <- as.numeric(
    names(K_distribution)
  )
  
  K_weights <- as.numeric(
    K_distribution
  )
  
  p <- pmin(
    pmax(p, 0),
    1
  )
  
  annual_risk <- numeric(
    length(p)
  )
  
  for (i in seq_along(K_values)) {
    
    annual_risk <- annual_risk +
      K_weights[i] *
      (
        1 -
          (1 - p)^K_values[i]
      )
  }
  
  annual_risk
}

calc_empirical_curve <- function(
  dose,
  fr,
  C,
  K_j
) {
  
  n_dose <- length(dose)
  n_pd <- nrow(pd_parameters)
  
  out <- calc_sd_drm(
    dose = rep(
      dose,
      each = n_pd
    ),
    
    fr = fr,
    C = C,
    alpha = alpha,
    beta = beta,
    
    Emax = rep(
      pd_parameters$Emax_per_day,
      times = n_dose
    ),
    
    EC50 = rep(
      pd_parameters$EC50_ug_mL,
      times = n_dose
    ),
    
    tfs = tfs
  )
  
  annual_risk_matrix <- matrix(
    annualise_risk(
      out$risk_total,
      K_j
    ),
    nrow = n_pd
  )
  
  status_matrix <- matrix(
    as.character(out$status),
    nrow = n_pd
  )
  
  list(
    mean_annual_risk =
      colMeans(annual_risk_matrix),
    
    status =
      apply(
        status_matrix,
        2,
        function(x) {
          names(
            which.max(
              table(x)
            )
          )
        }
      )
  )
}
```

# Scenario exploration functions

```{r scenario-exploration-functions}
make_annual_dose_response_data <- function(
  vary_by,
  values,
  fixed_fr = NULL,
  fixed_pMIC = NULL,
  n_person_years_curve = n_person_years,
  seed = annual_seed
) {
  
  set.seed(seed)
  
  F_j <- rnorm(
    n_person_years_curve,
    consumption_mean,
    consumption_se
  )
  
  F_j <- pmin(
    pmax(F_j, 0),
    1
  )
  
  K_j <- rbinom(
    n_person_years_curve,
    days_per_year,
    F_j
  )
  
  do.call(
    rbind,
    lapply(
      names(scenario_list),
      function(scn) {
        
        dose_vector <- scenario_list[[scn]]
        
        positive_dose <- dose_vector[
          dose_vector > 0 &
            is.finite(dose_vector)
        ]
        
        dose_grid <- 10^seq(
          log10(
            min(positive_dose)
          ),
          log10(
            max(positive_dose)
          ),
          length.out = 300
        )
        
        do.call(
          rbind,
          lapply(
            values,
            function(v) {
              
              fr_use <- if (
                vary_by == "fr"
              ) {
                v
              } else {
                fixed_fr
              }
              
              pMIC_use <- if (
                vary_by == "pMIC"
              ) {
                v
              } else {
                fixed_pMIC
              }
              
              out <- calc_empirical_curve(
                dose = dose_grid,
                fr = fr_use,
                C = (
                  pMIC_use / 100
                ) * MIC,
                K_j = K_j
              )
              
              data.frame(
                Scenario = scn,
                dose = dose_grid,
                value = v,
                
                group = if (
                  vary_by == "pMIC"
                ) {
                  paste0(
                    v,
                    "% MIC"
                  )
                } else {
                  paste0(
                    "fr = ",
                    v
                  )
                },
                
                mean_annual_risk =
                  out$mean_annual_risk,
                
                status =
                  out$status
              )
            }
          )
        )
      }
    )
  )
}

summarise_scenario_annual_risk <- function(
  vary_by,
  values,
  fixed_fr = NULL,
  fixed_pMIC = NULL,
  seed = annual_seed
) {
  
  do.call(
    rbind,
    lapply(
      names(scenario_list),
      function(scn) {
        
        dose_vector <- scenario_list[[scn]]
        
        do.call(
          rbind,
          lapply(
            values,
            function(v) {
              
              fr_use <- if (
                vary_by == "fr"
              ) {
                v
              } else {
                fixed_fr
              }
              
              pMIC_use <- if (
                vary_by == "pMIC"
              ) {
                v
              } else {
                fixed_pMIC
              }
              
              event_risk <- calc_sd_drm(
                dose = dose_vector,
                fr = fr_use,
                C = (
                  pMIC_use / 100
                ) * MIC,
                alpha = alpha,
                beta = beta,
                Emax = Emax_mc,
                EC50 = EC50_mc,
                tfs = tfs
              )
              
              annual_data <- simulate_annual_risk(
                daily_risk =
                  event_risk$risk_total,
                
                n_person_years =
                  n_person_years,
                
                consumption_mean =
                  consumption_mean,
                
                consumption_se =
                  consumption_se,
                
                days_per_year =
                  days_per_year,
                
                seed =
                  seed
              )
              
              annual_risk <-
                annual_data$annual_risk
              
              positive_annual_risk <-
                annual_risk[
                  annual_risk > 0
                ]
              
              conditional_quantiles <- quantile(
                positive_annual_risk,
                probs = c(
                  0.025,
                  0.125,
                  0.25,
                  0.50,
                  0.75,
                  0.875,
                  0.975
                ),
                names = FALSE
              )
              
              mean_less_risk <- mean(
                event_risk$risk_less_treatable
              )
              
              mean_more_risk <- mean(
                event_risk$risk_more_treatable
              )
              
              data.frame(
                Scenario = scn,
                value = v,
                
                probability_zero_annual_risk =
                  mean(
                    annual_risk == 0
                  ),
                
                mean_annual_risk =
                  mean(annual_risk),
                
                p025_positive_annual_risk =
                  conditional_quantiles[1],
                
                p125_positive_annual_risk =
                  conditional_quantiles[2],
                
                p25_positive_annual_risk =
                  conditional_quantiles[3],
                
                median_positive_annual_risk =
                  conditional_quantiles[4],
                
                p75_positive_annual_risk =
                  conditional_quantiles[5],
                
                p875_positive_annual_risk =
                  conditional_quantiles[6],
                
                p975_positive_annual_risk =
                  conditional_quantiles[7],
                
                mean_consumption_days =
                  mean(annual_data$K_j),
                
                outcome = ifelse(
                  mean_less_risk >=
                    mean_more_risk,
                  "Less likely treatable",
                  "More likely treatable"
                )
              )
            }
          )
        )
      }
    )
  )
}

prepare_annual_summary <- function(data) {
  
  data$Scenario <- factor(
    data$Scenario,
    levels = scenario_levels
  )
  
  data$outcome <- factor(
    data$outcome,
    levels = c(
      "More likely treatable",
      "Less likely treatable"
    )
  )
  
  data[
    order(
      data$Scenario,
      data$value
    ),
  ]
}

make_log10_annual_summary <- function(
  data,
  parameter_name
) {
  
  result <- data.frame(
    Scenario =
      as.character(data$Scenario),
    
    value =
      data$value,
    
    probability_zero_annual_risk =
      data$probability_zero_annual_risk,
    
    mean_consumption_days =
      data$mean_consumption_days,
    
    outcome =
      as.character(data$outcome),
    
    log10_mean_annual_risk =
      safe_log10(
        data$mean_annual_risk
      ),
    
    log10_p025_positive_annual_risk =
      safe_log10(
        data$p025_positive_annual_risk
      ),
    
    log10_p125_positive_annual_risk =
      safe_log10(
        data$p125_positive_annual_risk
      ),
    
    log10_p25_positive_annual_risk =
      safe_log10(
        data$p25_positive_annual_risk
      ),
    
    log10_median_positive_annual_risk =
      safe_log10(
        data$median_positive_annual_risk
      ),
    
    log10_p75_positive_annual_risk =
      safe_log10(
        data$p75_positive_annual_risk
      ),
    
    log10_p875_positive_annual_risk =
      safe_log10(
        data$p875_positive_annual_risk
      ),
    
    log10_p975_positive_annual_risk =
      safe_log10(
        data$p975_positive_annual_risk
      )
  )
  
  names(result)[
    names(result) == "value"
  ] <- parameter_name
  
  result
}

make_conditional_median_plot_data <- function(
  data,
  parameter_name
) {
  
  result <- data.frame(
    Scenario =
      as.character(data$Scenario),
    
    value =
      data$value,
    
    outcome =
      as.character(data$outcome),
    
    log10_median_annual_risk =
      safe_log10(
        data$median_positive_annual_risk
      ),
    
    log10_p025_annual_risk =
      safe_log10(
        data$p025_positive_annual_risk
      ),
    
    log10_p125_annual_risk =
      safe_log10(
        data$p125_positive_annual_risk
      ),
    
    log10_p25_annual_risk =
      safe_log10(
        data$p25_positive_annual_risk
      ),
    
    log10_p75_annual_risk =
      safe_log10(
        data$p75_positive_annual_risk
      ),
    
    log10_p875_annual_risk =
      safe_log10(
        data$p875_positive_annual_risk
      ),
    
    log10_p975_annual_risk =
      safe_log10(
        data$p975_positive_annual_risk
      )
  )
  
  names(result)[
    names(result) == "value"
  ] <- parameter_name
  
  result
}

save_annual_outputs <- function(
  data,
  parameter_name,
  summary_file,
  plot_data_file
) {
  
  write.csv(
    make_log10_annual_summary(
      data,
      parameter_name
    ),
    file.path(
      "results",
      summary_file
    ),
    row.names = FALSE
  )
  
  write.csv(
    make_conditional_median_plot_data(
      data,
      parameter_name
    ),
    file.path(
      "results",
      plot_data_file
    ),
    row.names = FALSE
  )
}

make_selected_annual_summary <- function(
  data,
  selected_value,
  scenario_exploration,
  selected_condition
) {
  
  x <- subset(
    data,
    value == selected_value
  )
  
  data.frame(
    Scenario_exploration =
      scenario_exploration,
    
    Selected_condition =
      selected_condition,
    
    Exposure_scenario =
      as.character(x$Scenario),
    
    Probability_zero_annual_risk =
      round(
        x$probability_zero_annual_risk,
        4
      ),
    
    Mean_consumption_days =
      round(
        x$mean_consumption_days,
        2
      ),
    
    log10_mean_annual_risk =
      round(
        safe_log10(
          x$mean_annual_risk
        ),
        3
      ),
    
    Dominant_outcome =
      as.character(x$outcome)
  )
}
```

# Scenario exploration 1: fixed resistant fraction and varying %MIC

```{r fixed-fr-varying-pmic-summary}
annual_risk_summary_pMIC <-
  prepare_annual_summary(
    summarise_scenario_annual_risk(
      vary_by = "pMIC",
      values = pMIC_seq,
      fixed_fr = fr_fixed
    )
  )
```

# Save results for varying %MIC

```{r save-results-pmic}
save_annual_outputs(
  data = annual_risk_summary_pMIC,
  parameter_name = "pMIC",
  
  summary_file =
    "annual_risk_summary_varying_pMIC.csv",
  
  plot_data_file = paste0(
    "conditional_median_annual_risk_",
    "plot_data_varying_pMIC.csv"
  )
)
```

# Plots for fixed resistant fraction and varying %MIC

```{r fixed-fr-varying-pmic-plots}
annual_dose_response_pMIC <-
  make_annual_dose_response_data(
    vary_by = "pMIC",
    values = pMIC_plot,
    fixed_fr = fr_fixed
  )

plot_dose_response_function(
  data = annual_dose_response_pMIC,
  scenario_levels = scenario_levels,
  colour_title = "% MIC doxycycline",
  
  subtitle_text = paste0(
    "Mean annual risk at each event dose; ",
    "resistant fraction fixed at ",
    fr_fixed
  ),
  
  file_name = paste0(
    "doxycycline_annual_dose_response_",
    "fixed_fr_varying_pMIC.png"
  )
)

plot_mean_annual_risk_function(
  data = annual_risk_summary_pMIC,
  scenario_levels = scenario_levels,
  x_title = "% MIC doxycycline",
  
  subtitle_text = paste0(
    "Resistant fraction fixed at ",
    fr_fixed
  ),
  
  file_name = paste0(
    "doxycycline_mean_annual_risk_",
    "fixed_fr_varying_pMIC.png"
  ),
  
  x_breaks = pMIC_breaks
)

plot_conditional_median_annual_risk_function(
  data = annual_risk_summary_pMIC,
  scenario_levels = scenario_levels,
  x_title = "% MIC doxycycline",
  
  subtitle_text = paste0(
    "Central 50%, 75%, and 95% Monte Carlo intervals; ",
    "resistant fraction fixed at ",
    fr_fixed
  ),
  
  file_name = paste0(
    "doxycycline_conditional_median_annual_risk_",
    "fixed_fr_varying_pMIC.png"
  ),
  
  x_breaks = pMIC_breaks
)
```

# Scenario exploration 2: fixed %MIC and varying resistant fraction

```{r fixed-pmic-varying-fr-summary}
annual_risk_summary_fr <-
  prepare_annual_summary(
    summarise_scenario_annual_risk(
      vary_by = "fr",
      values = fr_seq,
      fixed_pMIC = pMIC_fixed
    )
  )
```

# Save results for varying resistant fraction

```{r save-results-fr}
save_annual_outputs(
  data = annual_risk_summary_fr,
  parameter_name = "fr",
  
  summary_file =
    "annual_risk_summary_varying_resistant_fraction.csv",
  
  plot_data_file = paste0(
    "conditional_median_annual_risk_",
    "plot_data_varying_resistant_fraction.csv"
  )
)
```

# Plots for fixed %MIC and varying resistant fraction

```{r fixed-pmic-varying-fr-plots}
annual_dose_response_fr <-
  make_annual_dose_response_data(
    vary_by = "fr",
    values = fr_plot,
    fixed_pMIC = pMIC_fixed
  )

plot_dose_response_function(
  data = annual_dose_response_fr,
  scenario_levels = scenario_levels,
  colour_title = "Resistant fraction",
  
  subtitle_text = paste0(
    "Mean annual risk at each event dose; ",
    "doxycycline fixed at ",
    pMIC_fixed,
    "% MIC"
  ),
  
  file_name = paste0(
    "doxycycline_annual_dose_response_",
    "fixed_pMIC_varying_fr.png"
  )
)

plot_mean_annual_risk_function(
  data = annual_risk_summary_fr,
  scenario_levels = scenario_levels,
  x_title = "Resistant fraction",
  
  subtitle_text = paste0(
    "Doxycycline fixed at ",
    pMIC_fixed,
    "% MIC"
  ),
  
  file_name = paste0(
    "doxycycline_mean_annual_risk_",
    "fixed_pMIC_varying_fr.png"
  ),
  
  x_breaks = fr_breaks
)

plot_conditional_median_annual_risk_function(
  data = annual_risk_summary_fr,
  scenario_levels = scenario_levels,
  x_title = "Resistant fraction",
  
  subtitle_text = paste0(
    "Central 50%, 75%, and 95% Monte Carlo intervals; ",
    "doxycycline fixed at ",
    pMIC_fixed,
    "% MIC"
  ),
  
  file_name = paste0(
    "doxycycline_conditional_median_annual_risk_",
    "fixed_pMIC_varying_fr.png"
  ),
  
  x_breaks = fr_breaks
)
```

# Selected mean annual-risk estimates

```{r selected-estimates}
selected_pMIC <- 0
selected_fr <- 0

summary_fixed_fr <- make_selected_annual_summary(
  data = annual_risk_summary_pMIC,
  selected_value = selected_pMIC,
  
  scenario_exploration =
    "Fixed resistant fraction; varying %MIC",
  
  selected_condition = paste0(
    "fr = ",
    fr_fixed,
    "; pMIC = ",
    selected_pMIC,
    "% MIC"
  )
)

summary_fixed_pMIC <- make_selected_annual_summary(
  data = annual_risk_summary_fr,
  selected_value = selected_fr,
  
  scenario_exploration =
    "Fixed %MIC; varying resistant fraction",
  
  selected_condition = paste0(
    "pMIC = ",
    pMIC_fixed,
    "% MIC; fr = ",
    selected_fr
  )
)

selected_summary <- rbind(
  summary_fixed_fr,
  summary_fixed_pMIC
)

knitr::kable(
  selected_summary,
  caption =
    "Selected SD-DRM mean annual infection-risk estimates"
)

print(
  selected_summary,
  row.names = FALSE
)
```